---
title: "Import FPDS delta files"
output: html_document
editor_options: 
  markdown: 
    wrap: 72
---

# Whole database imports from monthly delta files to ETL.FPDSdelta

## 1) Setup

This should run very quickly, but you will need to take steps 1.A-1.C
\### 1.A) Add your USAspending-local folder if needed.

If this is your first time importing a file on this computer, copy and
paste two of the directory specifying lines and change both lines to
match your USAspending-local folder. The code chunk below looks like
this:

``` r
} else if(dir.exists("C:\\Users\\grego\\Repos\\USAspending-local\\")) {  
path\<-"C:\\Users\\grego\\Repos\\USAspending-local\\"  
```

### 1.B) Unzip the delta file and provide the directory name

**Every time**, update the deltadir variable to the month that next
needs to be imported. (Note you'll need to have already unzipped the
delta file to create this folder). The code chunk below looks like this:
`r deltadir\<-"FY(All)\_All_Contracts_Delta_20250906"`

### 1.C) Have access to the SQL login and password.

I recommend keeping it in a password manager.

```{r setup}
library(dplyr)
library(tidyr)
library(tidyverse)
library(csis360)
library(readr)
library(sqldf)
library(odbc)
library(askpass)
library(DBI)


if(!exists("login"))
  login<-askpass("Please enter the SQL login account")
if(!exists("pwd"))
  pwd<-askpass("Please enter the SQL account password")

if(dir.exists("C:\\Users\\grego\\Repositories\\USAspending-local\\")){
  path<-"C:\\Users\\grego\\Repositories\\USAspending-local\\"
} else if(dir.exists("D:\\Users\\Greg\\Repositories\\USAspending-local\\")) {
  path<-"D:\\Users\\Greg\\Repositories\\USAspending-local\\"
} else if(dir.exists("F:\\Users\\Greg\\Repositories\\USAspending-local\\")) {
  path<-"F:\\Users\\Greg\\Repositories\\USAspending-local\\"
} else if(dir.exists("C:\\Users\\grego\\Repos\\USAspending-local\\")) {
  path<-"C:\\Users\\grego\\Repos\\USAspending-local\\"
} else{
  stop("USAspending-local dir location unknown")
}
deltadir<-"FY(All)_All_Contracts_Delta_20251206"



file.list<-list.files(file.path(path,deltadir))
file.list<-file.list[substr(file.list,nchar(file.list)-3,nchar(file.list))==".rda"]
# EXEC ETL.SP_LastModifiedDateHistory
```

## Pull all entries that have the CTU

```{r import_to_r}

  ctu<-askpass("Please enter the contract_transaction_unique_key")


#This should be quite fast, roughly 5 minutes per file.
#Import


match<-NULL
for (f in 1:length(file.list)){
  load(file.path(path,deltadir,file.list[f]))
  if (is.null(match))
    match<-fd %>% filter(contract_transaction_unique_key==ctu)
  else 
    match<-rbind(match,
               fd %>% filter(contract_transaction_unique_key==ctu))
}
  

match
```
## Pull all deleted entries
```{r import_to_r}

  vmcon <- dbConnect(odbc(),
                     Driver = "SQL Server",
                     Server = "vmsqldiig.database.windows.net",
                     Database = "CSIS360",
                     UID = login,
                     PWD =pwd)
  
untransferred_sql<-paste(		"select top 100 fd.action_date_fiscal_year,",
	"isnull(fd.IsTransferred,0) as IsTransferred,",
	"fd.correction_delete_ind,",
	"fd.contract_transaction_unique_key,",
	"ctu.CSIStransactionID,",
	"fd.last_modified_date as source_last_modified_date,",
	"f.last_modified_date as target_last_modified_date",
"from ETL.FPDSdelta fd ",
	"left outer join contract.contract_transaction_unique_key ctu ",
	"on fd.contract_transaction_unique_key=ctu.contract_transaction_unique_key",
"left outer join contract.FPDS f",
	"on ctu.CSIStransactionID =f.CSIStransactionID",
"left outer join etl.FPDSdelta_support fds",
	"on ctu.contract_transaction_unique_key=fds.contract_transaction_unique_key")


#This will take a whie
untransferred<-dbGetQuery(vmcon,  untransferred_sql)

untransferred<-read.csv("../output/2026_12_06_faileddeletes.csv")
#This should be quite fast, roughly 5 minutes per file.
#Import


match<-NULL
for (f in 1:length(file.list)){
  load(file.path(path,deltadir,file.list[f]))
  if (!is.null(match))
    match<-fd %>% filter(contract_transaction_unique_key %in% untransferred$contract_transaction_unique_key)
  else
    match<-rbind(match,
               fd %>% filter(contract_transaction_unique_key %in% untransferred$contract_transaction_unique_key))
}
  

deleted
```
